### 2022-02-19 draft2.tex
Introduction of cc reworked and a few images added

\section{Introduction}
Forecasting has become increasingly important for private and public decision 
making. Because predictions cannot generally be free of uncertainty many 
thinkers prefer probabilistic forecasts over deterministic and probabilistic 
forecasts have become the norm in many fields including meteorology, economics,
finance and epidimiology.

Ensemble forecasting is the combination of several forecast models into one 
model which often yields better predicting power than the individual models
\cite{cramer2021evaluation} and has a solid station in dynamical weather 
prediction \cite{lewis2005roots}. In weather forecasting, individual members of 
an ensemble are generated from perturbations of initial conditions, see Baran 
section 2. Data \cite{baran2018combining} \cite{leutbecher2008ensemble}.
\cite{yamana2016superensemble}

A probabilistic forecasts may be represented in a variety of ways, a common way
being the form of a continuous or discrete probability distribution. Indeed much 
of the literature pertaining to scoring forecast models, combining models 
into ensemble forecasts and calibration pertain to methods for continuous
and/or discrete probability functions see for example 
\cite{gneiting2007probabilistic},
\cite{gneiting2013combining} \cite{baran2018combining}. Though such methods have
also been treated for samples, such as MCMC draws of a posterior predictive
distribution see for example \cite{krueger2016probabilistic}
and for quantile or interval type and expectile type forecasts 
\cite{taylor2021evaluating} \cite{bracher2021evaluating}.

Since 2013 the United States Center for Disease Control (CDC) has hosted an
annual competition for predicting the timing, peak and intensity of the year's 
flu season. In this competition, the CDC provides academic researchers with data
, relevant forecasting targets and other guidlines and the various teams produce
forecasts later evaluated against actual flu acticity. Details on the 
competition may be found on the CDC FluSight \cite{cdcflusight}.

The ReichLab of the University of Massachusettes Amherst and a 
CDC-funded Influenza Forecasting Centers of Excellence has been heavily 
involved in the CDC influenza competition and disease outbreak research.
In their work they have provided insight into the usefulness of collaboration 
and ensemble modeling in disease outbreaks, see for example 
\cite{reich2019collaborative}, \cite{reich2019accuracy} and 
\cite{mcandrew2019adaptively}.

In March 2020, at the onset of the COVID-19 pandemic, the Reich Lab in
association with others founded the COVID-19 Forecast Hub. Every week relevant
pandemic data is provided to several teams who build probabilistic models for
forecasting cases, hospitalizations and deaths due to COVID-19 for various 
regions of different levels with predictions of days, weeks and months ahead.
These forecasts are combined into a single ensemble forecast. The model data,
forecasts and the ensemble are passed along to the CDC which uses in official
communication \cite{Cramer2021-hub-dataset}.

Though similar to the forecasting in the influenza competition, the format of 
the COVID-19 Forecast Hub has some key distinctions. For one, the Forecast Hub
has as a key component the construction of a multi-model ensemble forecast 
whereas that has not as yet been the case in the influenza competition. Another 
key difference is the format which the individual forecasts are submitted. In 
the influenza forecast, a set of discretized bins on a numeric scale were set
for each target and participating researchers were asked to submit their 
probabilistic predictions as a list of probabilities in for each bin essentially
making each prediction a bounded pmf \cite{mcgowan2019collaborative}.
Due to the initial lack of understanding of COVID-19 and the time pressure of
creating forecasts, the forecasts were requested as the predictive mean and 
predictive intervals for various nominal levels depending on the target to be
predicted \cite{bracher2021evaluating}.

Though the representation of a set of intervals can bring added flexibility, 
there is also some loss in what can be said about a forecast. This will be 
discussed further herein. In this paper we propose as an alternative to 
requesting probabilistic forecasts in binned, quantiled, sampled or common 
distribution format requesting forecasts in a format conducive to discrete 
mixture distribution. Discussed will be advantages and disadvantages of
a forecast of this form.

In the next section, popular probabilistic forecast representation types
will be reviewed included standard pieces of definitions. These include 
continuous/discrete probability functions, discretized bin distributions,
samples from a distribution and quantile or interval distributions. In the 
following section common methods for scoring forecasts, constructing multi-model
ensembles will be reviewed and compared for the different representation types. 
In the same section, other practical considerations for a forecast hub will be
discussed.

@@

@@

\begin{table}[h!]
\centering
 \begin{tabular}{||c c c c||} 
 \hline
 Family & param1 & param2 & param3 \\ [0.5ex] 
 \hline\hline
 normal & 0 & 1 & NA \\ 
 \hline
 \end{tabular}
\end{table}

@@

\begin{table}[h!]
\centering
 \begin{tabular}{|c|c|} 
 \hline
    bin & prob \\ \hline
    -6.5 & 3.75e-11 \\
    -6.4 & 7.11e-11 \\
    -6.3 & 1.33e-10 \\
    ...  & ... \\
    6.3 & 7.11e-11 \\
    6.4 & 3.75e-11 \\
    6.5 & 1.96e-11 \\
 \hline
 \end{tabular}
\end{table}

@@

\begin{table}[h!]
\centering
 \begin{tabular}{|c|c|} 
 \hline
    quantile & value \\ \hline
    0.01 & -2.326 \\
    0.025 & -1.96 \\
    0.05 & -1.645 \\
    0.01 & -1.282 \\
    ...  & ... \\
    0.9 & 1.282 \\
    0.95 & 1.645 \\
    0.975 & 1.96 \\
    0.99 & 2.326 \\
 \hline
 \end{tabular}
\end{table}

>>>>

\section{Introduction}
Predicting the outcomes of prospective events is the object of much scientific
inquiry and the basis for many decisions both public and private. Because 
predictions of the future can almost never be precise, it is usually desirable
that a level of uncertainty be attached to any prediction. In recent years, it
has become increasingly desirable that forecasts be probabilistic in order to 
account for uncertainty in predicted quantities or events 
\cite{gneiting2014probabilistic}. Specific problems for 
which probabilistic forecasts is used include weather forecasting
\cite{baran2018combining}, economics \cite{groen2013real} and disease outbreaks
\cite{yamana2016superensemble}.

A probabilistic forecast is a probability distribution that may be represented
in a few different formats. A common representation is either a continuous or
discrete parametric distribution function (a pdf or pmf). Much of the literature 
on calibration, sharpness and scoring of a forecast pertains to parametric 
distribution forecasts (see for example 
\cite{gneiting2007probabilistic},
\cite{gneiting2013combining} \cite{baran2018combining}).
Other common representations include samples \cite{krueger2016probabilistic}, 
discretized bins 
and quantile or interval type forecasts \cite{taylor2021evaluating} 
\cite{bracher2021evaluating}. Each representation may be more or less 
appropriate than the others for a given problem but knowing how to interpret, 
score and contruct ensemble forecasts for a selected representation is a must 
when multiple forecasts of the same event are involved.

Two projects on forecasting disease outbreaks for which many separate forecasts 
are used include the United States Center for Disease Control (CDC) hosted
annual competition for forecasting influenza \cite{cdcflusight}
and the COVID-19 Forecast Hub which has 
continuously operated since the start of the COVID-19 pandemic in the US in 
early 2020 \cite{Cramer2021-hub-dataset}.

\subsection{CDC Influenza Like Illness}

Since the 2013-14 flu season, the CDC has hosted an annual competition for 
predicting the timing, peak and intensity of the year's flu season. Forecasts
for the different targets also include predictions for one, two, three and 
four weeks ahead of the prediction time. National flu data is weekly provided 
to outside academic teams who use that data to construct forecasts however they
will, but the forecast predictions are always submitted in a discretized bin
format. Here the binning scheme was on a bounded numeric scale and the 
prediction of a specific target was a set of probabilities assigned to each bin
\cite{mcgowan2019collaborative}.
These forecasts are then evaluated against actual flu activity, and at 
the end of the season a winning team is declared \cite{cdcflusight}.

This competition has provided the CDC, competing teams and other interested
parties a chance to collaborate and improve forecasting each season. One 
proposed way to enhance prediction has been to aggregate the various by team
forecasts into a multi-model ensemble forecast \cite{mcgowan2019collaborative}
\cite{mcandrew2019adaptively} \cite{reich2019accuracy}.

A multi-model ensemble forecast is combination of several component forecast 
models into one model which often yields better predicting power than the 
individual models \cite{cramer2021evaluation} (and did lead to better average 
results with flu competition forecasts in \cite{reich2019accuracy}.) 
In the COVID-19 Forecast Hub, construction of ensemble forecasts is a main
priority.

% Ensemble building has a history in dynamical weather 
% prediction \cite{lewis2005roots}. In weather forecasting, individual members of 
% an ensemble are generated from perturbations of initial conditions, see Baran 
% section 2. Data \cite{baran2018combining} \cite{leutbecher2008ensemble}.
% \cite{yamana2016superensemble}

\subsection{COVID-19 Forecast Hub}
In March 2020, at the onset of the COVID-19 pandemic, the Reich Lab in
association with others founded the COVID-19 Forecast Hub. Borrowing on work and 
ideas from the CDC annual competition, the Forecast Hub is a central site in 
which dozens of academic teams collaborate to forecast the ongoing COVID-19 
pandemic.
Every week relevant
pandemic data is provided to these teams who build probabilistic models for 
their use in
forecasting cases, hospitalizations and deaths due to COVID-19. Forecasts are 
made on the US county,
state and national level with predictions for days, weeks and months ahead.
These forecasts are combined into a single ensemble forecast. The model data,
forecasts and the ensemble are passed along to the CDC for its use in official
communication \cite{Cramer2021-hub-dataset}.

Though similar to the forecasting in the influenza competition, the format of 
the COVID-19 Forecast Hub has some key distinctions. For one this project has
been operating continuously since it first began, so forecasts have been made
for over one-hundred straight weeks. 
As well due to the initial lack of understanding of COVID-19 and the time 
pressure of
creating forecasts, rather than requesting forecasts as binned probabilities
the forecasts were requested as the predictive mean and 
predictive intervals for various nominal levels depending on the target to be
predicted \cite{bracher2021evaluating}. Collecting forecasts in this quantile
or interval type brings with it differences in how to score, construct ensemble
models and store the forecasts along with other considerations.

\begin{figure}[htbp]
\centerline{\includegraphics[scale=.12]{8_30_21_cvd_deaths.jpeg}}
\caption{Figure this out later}
\label{fig}
\end{figure}

\subsection{...}
In the context of collaborative forecasting like that of the CDC flu competition
or the COVID-19 Forecast Hub, discretized bins and quantile forecasts have 
proven useful and effective. Yet both types come with their drawbacks. Data 
storage for instance might be a concern if many bins are allowed for a 
prediction. And scoring methods are limited if forecasts are made of quantiles.

In this paper, we propose the use of discrete mixture distributions as a means 
of forecasting for projects similar to the CDC flu competition or the COVID-19
Forecast Hub. In section 2, popular probabilistic forecast representation 
types will be defined and reviewed. Methods of scoring, storing, building 
ensembles and other aspects will be reviewed and compared for each 
representation type.
Section 3 is a retrospective study of the CDC flu competition and COVID-19 
forecast predictions and an attempt to assess whether any predictions 
approximately come from well known continuous distributions.

@@

\begin{figure}[htbp]
\centerline{\includegraphics[scale=.4]{dens_comp.png}}
\caption{Figure this out later}
\label{fig}
\end{figure}

\begin{figure}[htbp]
\centerline{\includegraphics[scale=.4]{cum_comp.png}}
\caption{Figure this out later}
\label{fig}
\end{figure}

@@

\begin{table}[h!]
\centering
 \begin{tabular}{||c c c c||} 
 \hline
 Family & param1 & param2 & param3 \\ [0.5ex] 
 \hline\hline
 lnorm & 1 & 0.4 & NA \\ 
 \hline
 \end{tabular}
\end{table}

@@

\begin{table}[h!]
\centering
 \begin{tabular}{|c|c|} 
 \hline
    bin & prob \\ \hline
    ... & ... \\
    {[1.4,\;1.6)} & 0.04414 \\
    {[1.6,\;1.8)} & 0.05896 \\
    {[1.8,\;2.0)} & 0.07032 \\
    {[2.0,\;2.2)} & 0.07172 \\
    {[2.2,\;2.4)} & 0.07955 \\
    ... & ... \\
 \hline
 \end{tabular}
\end{table}

@@

\begin{table}[h!]
\centering
 \begin{tabular}{|c|c|} 
 \hline
    quantile & value \\ \hline
    0.01 & 1.07137 \\
    0.025 & 1.2404 \\
    0.05 & 1.40689 \\
    0.01 & 1.62675 \\
    ...  & ... \\
    0.9 & 4.50667 \\
    0.95 & 5.18328 \\
    0.975 & 5.82391 \\
    0.99 & 6.58783 \\
 \hline
 \end{tabular}
\end{table}
